Nail: A practical interface generator for binary formats

### 2014-04-01 paper/design.tex
Rebased some changes

In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
by the combinator of conventional Parser Expression Grammars they replace.
\paragraph{Fundamental parsers}
The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
Examples: \texttt{uint4}, \texttt{int32}

Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
\texttt{int32| [1,255,512]}

>>>>

In the following, we present the combinators in Nail grouped by the combinators of conventional Parser Expression Grammars they replace.
\paragraph{Fundamental parsers}
The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
Examples: \texttt{uint4}, \texttt{int32}

Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
element of a set of allowed values. \\
Examples: \texttt{uint8 | 1..3} (accepts an 8 bit unsigned
integer between 1 and 3 inclusive) ,\\
 \texttt{uint16| ..512} (an unsigned 16-bit integer less or
equal than 512),\\
\texttt{int32| [1,255,512]} (a signed 32 bit integer with value 1, 255 or 512).

### 2014-04-01 paper/intro.tex
split off relwk
Rebased some changes

\section{Background}

@@

Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
  differently, so the code signature sees different bytes at a virtual address than the executable
  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
  files. Signatures are verified with a Java program, but the program is extracted with a C program.
The java program interprets all fields as signed, whereas the C program treats them as unsigned,
allowing one to replace files in a signed archive.}
\cite{saurik-masterkey} code signing and even the
X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
runtime environment or static analysis guarantee absolute memory safety.

@@

 Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
 While previous parser generators could also be used to write grammars for binary
 protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
 inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
 of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
 library. Hammer implements grammars as language-integrated parser combinators, an approach
 popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
 knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
 top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).

>>>>

\section{Introduction}

@@

Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
  differently, so the code signature sees different bytes at a virtual address than the executable
  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
  files. Signatures are verified with a Java program, but the program is extracted with a C program.
The java program interprets all fields as signed, whereas the C program treats them as unsigned,
allowing one to replace files in a signed archive.}
\cite{saurik-masterkey} code signing and even the
X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
runtime environment or static analysis guarantee memory safety.

@@

 Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
 While previous parser generators could also be used to write grammars for binary
 protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
 inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
 of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
 library. Hammer implements grammars as language-integrated parser combinators, an approach
 popularised by Parsec for Haskell\cite{LeijenMeijer:parsec}. The parser combinator style (to our
 knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
 top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).

### 2014-04-01 paper/relwk.tex
Rebased some changes

Generating parsers and generators from an executable specification is the core concept of Interface
Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
generators work very well for existing grammars, they do not allow full control over the format of
the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
implements a C like structure model enhanced with data-dependent length fields and constraints.

>>>>

Generating parsers and generators from an executable specification is the core concept of Interface
Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
generators work very well for existing grammars, they do not allow full control over the format of
the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
however PacketTypes works only as a parser, not as an output generator and does not support the expressive power of PEGs, but rather
implements a C like structure model enhanced with data-dependent length fields and constraints.
